AtmoSwing (v1.4): Analogue Technique Model for Statistical weather forecastING

### 2016-05-27 AtmoSwing - Geoscientific Model Development/template.tex
Addition of the variables exploration analysis.

@@

\citet{Horton2012} identified that the analogy on the atmospheric circulation may not be comprehensive in an alpine environment, and that a significant gain was obtained by considering 4 atmospheric levels (1000, 850, 700, and 500~hPa at 12 or 24~h) instead of 2 (Table \ref{table:R6}). The number of analogues ($N_{1}$) was not provided globally, as it was calibrated for every subregion. An equivalent structure was also optimized by means of generic algorithms (PC-4Zgo), which is the topic of another paper \citep{Horton2016b}.

@@

\citet{Horton2012} also integrated an analogy on the moisture flux, after the circulation analogy on the 4 atmospheric levels (Table \ref{table:R7}). The number of analogues ($N_{1}$, $N_{2}$) were calibrated for every subregion. Similarly, an equivalent structure was optimized \citep[PC-4Zgo-2MIgo,][]{Horton2016b}.

\citet{BenDaoud2010} applied the AM in the context of large floodplains (Sa\^{o}ne, Seine). He evaluated several variables usually used in weather forecasting, amongst them the air temperature and the vertical velocity that presented a potential interest. The temperature (at the nearest grid point, on the 925~hPa \& 36~h and 600~hPa \& 12~h levels) could replace the rigid calendar preselection of $\pm$ 60 days around the target date by a more dynamic screening of similar situations in terms of air masses. The seasonal effect is indeed also present in the temperature data. The number of preselected dates is equivalent to the number of days one would have chosen with the calendar approach, and thus depends on the archive size: $N_{0} = 60 \cdot 2 \cdot n_{a}$ where $n_{a}$ is the number of years in the calibration period. \citet{BenDaoud2010} also reconsidered the parameters of the moisture index and ended up with both 925~hPa and 700~hPa levels, at 12~h and 24~h, or at every time step (6~h) between 06~h and 30~h (Table \ref{table:R4b}).

@@

@@

\begin{figure}[t]
	\includegraphics[width=8.3cm]{figures/fig03.pdf}
	\caption{Performance scores, for the Swiss Alps, for different AM variants on multiple levels of analogy \citep{Horton2012a}.}
	\label{figure:scores_2}
\end{figure}

>>>>

\section{Data and case study description}
\label{sec:data_and_case_study}

\subsection{Case study description}
\label{sec:case_study}

Some results will be illustrated throughout this article for the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{figure:map}). Even though the region is rather small (5524~km$^{2}$), the meteorological influences related to extreme weather conditions varies substantially within it \citep[see][]{Horton2012}. The gauging stations in the catchment were thus aggregated into 10~subregions (Fig.\ \ref{figure:map}): (1) Swiss Chablais, (2) Trient Valley, (3) West Bernese Alps, (4) Lower Rhone Valley, (5) Left side valleys, (6) Southern ridges, (7) Upper Rhone Valley, (8) Southeast ridges, (9) East Bernese Alps, and (10) Conches Valley.

@@

Every variable of the NCEP/NCAR reanalysis was systematically assessed for each of the identified subregions of the study area (Sect. \ref{sec:case_study}). The present assessment considers every subregion separately and optimizes the spatial window and the number of analogues for every combination of variables, pressure levels, times of observation, and subregions (using the sequential calibration tool implemented in AtmoSwing, see Sect. \ref{sec:atmoswing-calibration}). Even though the best predictors may vary from one place to another due to different meteorological influences (see Fig. \ref{figure:variable_exploration_Chablais} for the 30 best predictors for the Chablais region, and Fig. \ref{figure:variable_exploration_SE_crests} for the Southeast ridges), 2 main conclusions can be drawn from the NCEP/NCAR reanalysis dataset for the area of interest, that meet those from \citet{Bontron2004}: (1) the variables describing the atmospheric circulation (pressure fields or geopotential heights) are the ones with the highest prediction capacity (the precipitation rate is even rarely selected in the 30 best predictors), and (2) these are better when compared by means of the S1 criteria (asterisk in Fig. \ref{figure:variable_exploration_Chablais} and \ref{figure:variable_exploration_SE_crests}) instead of the RMSE. The reason for this latter point is the better representation of the atmospheric circulation information when considering the shape of a pressure field rather than its values.

\citet{Horton2012} identified that the analogy on the atmospheric circulation as defined by \citet{Bontron2004} may not be comprehensive in an alpine environment, and that considering 4 atmospheric levels (1000, 850, 700, and 500~hPa at 12 or 24~h) instead of 2 could improve the performance (Table \ref{table:R6} -- the number of analogues $N_{1}$ was not provided globally, as it was calibrated for every subregion). The change in performance scores is illustrated in Fig. \ref{figure:scores_1} for the different subregions of the study area. An equivalent structure was also optimized by means of generic algorithms (PC-4Zgo), which is the topic of another paper \citep{Horton2016b}. As these new parametrizations provide performance gains for alpine catchments under varying meteorological influences (Fig. \ref{figure:scores_1}), they are recommended for operational use.

@@

\citet{Horton2012} also integrated an analogy on the moisture flux, after the circulation analogy on the 4 atmospheric levels (Table \ref{table:R7}). The number of analogues ($N_{1}$, $N_{2}$) were calibrated for every subregion. Similarly, an equivalent structure was optimized \citep[PC-4Zgo-2MIgo,][]{Horton2016b}. The change in the performance scores, for the method with 2 levels of analogy, is illustrated in Fig. \ref{figure:scores_2} for the study area. As previously, we would recommend using these new parametrizations for alpine catchments.

\citet{BenDaoud2010} applied the AM in the context of large floodplains (Sa\^{o}ne, Seine). He evaluated several variables usually used in weather forecasting, amongst them the air temperature and the vertical velocity that presented a potential interest. The temperature (at the nearest grid point, on the 925~hPa \& 36~h and 600~hPa \& 12~h levels) could replace the rigid calendar preselection of $\pm$ 60 days around the target date by a more dynamic screening of similar situations in terms of air masses. The seasonal effect is indeed also present in the temperature data. The number of preselected dates is equivalent to the number of days one would have chosen with the calendar approach, and thus depends on the archive size: $N_{0} = 60 \cdot 2 \cdot n_{a}$ where $n_{a}$ is the number of years in the calibration period. \citet{BenDaoud2010} also reconsidered the parameters of the moisture index and ended up with both 925~hPa and 700~hPa levels, at 12~h and 24~h, or at every time step (6~h) between 06~h and 30~h (Table \ref{table:R4b}). The performance of this parameterization for our case study is presented in Fig. \ref{figure:scores_2}.

@@

\begin{figure}[t]
	\includegraphics[width=8.3cm]{figures/map.pdf}
	\caption{Location of the alpine Rh\^{o}ne catchment in Switzerland. (source: Swisstopo)}
	\label{figure:map}
\end{figure}

\begin{figure}[t]
	\includegraphics[width=7.5cm]{figures/variable_exploration_Chablais.pdf}
	\caption{Performance score (CRPSS) of the 30 best variables from the NCEP/NCAR reanalysis dataset, when considered separately (no combination), for the Chablais region (Sect. \ref{sec:case_study}). The analogy criteria is S1 when there is an asterisk next to the variable name, and RMSE otherwise. Color illustrates the variable type: green = atmospheric circulation, blue = moisture, orange = temperature, yellow = radiation, purple = vertical velocity, and gray = other. SLP stands for sea level pressure, and Z for geopotential height}
	\label{figure:variable_exploration_Chablais}
\end{figure}

\begin{figure}[t]
	\includegraphics[width=7.5cm]{figures/variable_exploration_SE_crests.pdf}
	\caption{Same as Fig. \ref{figure:variable_exploration_SE_crests}, but for the southeast ridges.}
	\label{figure:variable_exploration_SE_crests}
\end{figure}

@@

\begin{figure}[t]
	\includegraphics[width=8.3cm]{figures/fig03.pdf}
	\caption{Performance scores, for the Swiss Alps, for different AM variants on multiple levels of analogy \citep{Horton2012a}. Same conventions as Fig. \ref{figure:scores_2}.}
	\label{figure:scores_2}
\end{figure}
